Nuclei Segmentation in Histopathology images using Deep Neural Networks

### 2016-11-07 Paper.tex

%% Data augmentation and Hyper parameter training: Training scheme
%% Check number of images and number of transformations
To train our models, as the number of available annotated is scarse, we 
used a great number of transformation for the data augmentation. From 
a original size of 33 annotated images, adding flips, rotations, bluriness 
and random elastic deformations enabled us to have more then 
$400 000$ training images. 
We also try out several hyperparameter configurations: the learning rate and 
momentum for the stochastic gradient descent, the weight decay value.  
In practice, we found that hyperparameters tuning didn't influence the 
scores much, the exepction being the learning rate. If the learning rate 
was not of the right magnitude the given network did not seem to learn. 
We also experienced with different 
initialization value and if possible, we also considered pretrained layers. 
Using pretrained layers made learning more efficient and made score 
values more robust.

@@

\section{Different architectures/Results}
\label{sec:results}
We experience with 4 known arhitectures in semantic segmentation, one 
will be named BaochuanNet, Fully Convolutionnal Net (FCN), DeconvNet, 
UNet. The most basic architecture, BaochuanNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
archicture has the advantage of re-using a classical deep learning 
architecture with added on upsampling layer and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layer as one intends to learn 
the proper upsampling threw repeated deconvolution and convolution 
layers. This model can also be fine tuned. The final architecture that we tried is the UNet. This network combines both contribution of the above networks, the UNet has a similar architecture to the DeconvNet with skip layers.
\section{Result}
\label{sec:result}

\begin{tabular}{|c|c|c|c|c|c|}
 & UNet & Baochuan & DeconvNet & FCN & Ensemble \\
%\hrule
Accuracy          &       0.89 &       0.94 & 0.97\\
IU   &       0.70 &       0.76 &     0.86 \\
Recall                     &       0.75 &       0.74  &     0.86 &&\\
Precision                  &       0.64 &       0.74 &     0.86 &&\\
F1                         &       0.69 &       0.74 &     0.86 &&  \\
TP             &       0.75 &       0.74 &     0.86 && \\
TN             &       0.92 &       0.96 &     0.98 &&\\
Performance                &       0.84 &       0.85 &     0.92 &&\\
Pixel error                &       0.11 &       0.06 &     0.03 &&\\
\end{tabular}

\section{Discussion}
\label{sec:discussion}
What's next?

@@

% Below is an example of how to insert images. Delete the ``\vspace'' line,
% uncomment the preceding line ``\centerline...'' and replace ``imageX.ps''
% with a suitable PostScript file name.
% -------------------------------------------------------------------------
\begin{figure}[htb]

\begin{minipage}[b]{1.0\linewidth}
  \centering
  \centerline{\includegraphics[width=8.5cm]{RGB_1}}
%  \vspace{2.0cm}
  \centerline{Patient 498959}\medskip
\end{minipage}
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_1}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 498959}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_1}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_2}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 581910}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_2}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_3}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 581910}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_3}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
\caption{Random annotated samples from the dataset}
\label{fig:annotation}
%
\end{figure}

>>>>

%% Data augmentation and Hyper parameter training: Training scheme
%% Check number of images and number of transformations
To train our models, as the number of available annotated is scarse, we 
used a great number of transformation for the data augmentation. From 
a original size of 33 annotated images, adding flips, rotations, bluriness 
and random elastic deformations enabled us to have more then 
$400 000$ training images. 
We also try out several hyperparameter configurations: the learning rate and 
momentum for the stochastic gradient descent, the weight decay value.  
In practice, we found that hyperparameters tuning didn't influence the 
scores much, the exepction being the learning rate. If the learning rate 
was not of the right magnitude the given network did not seem to learn. 
We therefore fixed the momentum to be 0.9 and the weight decay to be 
$5.10^{-5}$ for all of the experiences, the learning rate was tunned 
according to the model choosen.
We also experienced with different 
initialization value and if possible, we also considered pretrained layers. 
Using pretrained layers made learning more efficient and made score 
values more robust.

@@

\section{Different architectures/Results}
\label{sec:results}
We experience with 4 known arhitectures in semantic segmentation, one 
will be named BaochuanNet, Fully Convolutionnal Net (FCN), DeconvNet, 
UNet. The most basic architecture, BaochuanNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
archicture has the advantage of re-using a classical deep learning 
architecture with added on upsampling layer and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture, for a more detailed explaination please refer to \cite{long2015fcn}. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layer as one intends to learn 
the proper upsampling threw repeated deconvolution and convolution 
layers. This model can also be fine tuned, for more information please 
refer to the original paper \cite{noh2015learning}. We did not
ensemble classifier as suggested in 
\cite{noh2015learning} as our FCN and DeconvNet classifier achieve very similar scores.
% The final architecture that we tried is the UNet. This network combines 
%both contribution of the above networks, the UNet has a similar 
%architecture to the DeconvNet with skip layers.
\section{Result and discussion}
\label{sec:result}
We conducted our experiments on GPUs via the caffe framework, \cite{jia2014caffe}. We present our results in table \ref{tab:res} where we averaged the metrics other all left out patients. We also added, for illustration purposes \ref{fig:prediction}, image prediction and associated probability maps to evaluate some differences between the classifiers. We notice that the simpler net, Baochuan Net seems to have learnt simple rules based on color information. As we can not notice by the "holes" in the probability map when the interior of the cell is brighter. On the contrary, deeper nets as FCN and DeconvNet have learnt deeper features and seem capable of recognizing whole cells. The small noticable differences in \ref{fig:prediction} are due to a random crop, this results in a slightly shiffted image for each one of the predictions.
\begin{table}
\begin{tabular}{|c|c|c|c|}
\hline
  & Baochuan & DeconvNet & FCN \\
 \hline
Accuracy  &       0.936 & 0.968 &0.958  \\
IU   &    0.759 &     0.856 & 0.857 \\
Recall     &       0.744  &     0.856 & 0.855 \\
Precision   &       0.741 &     0.858 & 0.878\\
F1    &       0.742 &     0.856 & 0.866 \\
TP       &       0.744 &     0.856 & 0.855  \\
TN      &       0.963 &     0.981 &  0.977\\
Performance    &       0.853 &     0.918 & 0.916\\
Pixel error  &  0.063 &     0.032 & 0.042\\
\hline
\end{tabular}
\caption{\label{tab:res}Results}
\label{tab:res}
\end{table}

%%with Unet
%\begin{tabular}{|c|c|c|c|c|c|}
% & UNet & Baochuan & DeconvNet & FCN & Ensemble \\
%\hrule
%Accuracy &       0.89 &       0.936 & 0.968 &0.958 & \\
%IU   &       0.70 &       0.759 &     0.856 & 0.857 &\\
%Recall    &       0.75 &       0.744  &     0.856 & 0.855 &\\
%Precision    &       0.64 &       0.741 &     0.858 & 0.878&\\
%F1       &       0.69 &       0.742 &     0.856 & 0.866&  \\
%TP             &       0.75 &       0.744 &     0.856 & 0.855 & \\
%TN             &       0.92 &       0.963 &     0.981 &  0.977&\\
%Performance    &       0.84 &       0.853 &     0.918 & 0.916&\\
%Pixel error  &       0.11 &       0.063 &     0.032 & 0.042&\\
%\end{tabular}

@@

% Below is an example of how to insert images. Delete the ``\vspace'' line,
% uncomment the preceding line ``\centerline...'' and replace ``imageX.ps''
% with a suitable PostScript file name.
% -------------------------------------------------------------------------
\begin{figure}[htb]
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_1}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 498959}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_1}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_2}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 581910}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_2}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_3}}
%  \vspace{1.5cm}
  \centerline{Another one from patient 581910}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_3}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
\caption{Random annotated samples from the dataset}
\label{fig:annotation}
%
\end{figure}

\begin{figure}[htb]
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{RGB_pred}}
%  \vspace{1.5cm}
  \centerline{Original RGB sample}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_pred}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}

\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{BaochuanB}}
%  \vspace{1.5cm}
  \centerline{BaochuanNet}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{BaochuanP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}

%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{DeconvNetB}}
%  \vspace{1.5cm}
  \centerline{DeconvNet}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{DeconvNetP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{FCNB}}
%  \vspace{1.5cm}
  \centerline{FCN}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{FCNP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}
%
\caption{Prediction via different classifiers of a random sample on the left out patient: 572123}
\label{fig:prediction}
%
\end{figure}
